---
title: "Mapping"
author: "Maggie Sha"
date: "2020/10/27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate, warn.conflicts = FALSE)
library(dplyr)
library(magrittr)

library(usmap)
library(maps)
```


## Introduction
This document works on the data from FEMA, and shows every step I have taken to acquire, clean, organize and map the data.

## Data Source
I am using a publically available dataset from FEMA.

  * The data is from https://www.fema.gov/openfema-data-page/public-assistance-funded-projects-details-v1
  * In this report, I only focus on Hurricane data from year 2009 to 2018.

## Data Cleaning and Organization
```{r cars}
# import data and filter hurricane data from 2009 to 2018
data <- read.csv("PublicAssistanceFundedProjectsDetails.csv",header=TRUE)
data <- data %>% filter(incidentType == "Hurricane") 
data$declarationDate <- ymd_hms(data$declarationDate)
data <- data %>% filter(2009 <= year(data$declarationDate)) 
data <- data %>% filter(year(data$declarationDate) <= 2018)

#unique(data$state)
 # state <- c("Alabama","Texas","Virgin Islands of the U.S.","North Carolina","Massachusetts","Puerto Rico",
 #            "New York","Virginia","New Hampshire","Maryland","Delaware","West Virginia","Louisiana","Florida",
 #            "New Jersey","Vermont","Connecticut","Pennsylvania","Rhode Island","Maine","District of Columbia",
 #            "Mississippi","Ohio","Georgia","South Carolina","American Samoa","Hawaii")

myState <- tolower(unique(data$state))
county <- tolower(unique(data$county))

MainStates <- map_data("state",myState)
AllCounty <- map_data("county",myState)

# length(unique(MainStates$region))
# [1] 23

# length(unique(AllCounty$subregion))
# [1] 914
```
There are 23 unique States and 914 counties in my hurricane data.


This is how the hurricane data looks like:
```{r}
head(data,n=10)
```


There are many attributes that I am not going to use or compare. 
I will only keep several columns and do future data cleaning from here.

```{r}
# select columns I need 
cTable <- data%>%select("declarationDate","county","countyCode","state","projectSize","projectAmount","federalShareObligated","totalObligated")

# get a new table contains projectTotal
sumProject <- cTable %>% group_by(countyCode) %>% dplyr::summarise(projectTotal = sum(projectAmount))

# merge cTable and projectTotal
cTable <- merge(sumProject,cTable,by = "countyCode")

cTable %<>% dplyr::rename(region=state, subregion=county)
cTable$region %<>% tolower()
cTable$subregion %<>% tolower()

# correct those counties' names which contain " (city)"
cTable$subregion <- gsub("\\ \\(city\\)","",cTable$subregion)


# join the table with Allcounty to match latitude and longitude
# filter out miss-matched rows
cTable <- left_join(cTable,AllCounty,by = "subregion")
cTable %<>% filter(long != "NA")

# find which counties have project amount greater than 100 million.
cTable %<>% mutate(`projectAmount > 100000000` = ifelse(cTable$projectTotal > 100000000, "True", "False"))


# length(unique(cTable$region.x))
# [1] 22

# length(unique(cTable$subregion))
# [1] 525
```

Then I am done with the data cleaning and organization part.
Now the cTable has 22 distinct States and 525 counties.
This is how cTable looks like:
```{r}
head(cTable,n=10)
write.csv(cTable,"D:\\MA615\\Mapping-pro\\cTable.csv", row.names = FALSE)
```


## Mapping

First I want to see the distribution of project size (either "large" or "small")
```{r}
ggplot() +
  geom_polygon(data=AllCounty, aes(x=long, y=lat, group=group),
               color="gray", fill="white", size = .1 ) +
  geom_polygon(data = cTable, aes(x = long, y = lat, group = group, 
                                  fill = `projectSize`),
                color = "grey", size = 0.2, alpha = 1.6) +
  geom_polygon(data = MainStates, aes(x = long, y = lat, group = group),
               color="black", fill="white", size = 0.2, alpha = 0.3) +
  scale_fill_manual(values = c("gold", "blue"))+ 
  ggtitle("Project Size") 

```


Then I map the project total amount grouped by county too see which exceed 100 million.
```{r}
  # use ggplot 
  ggplot() +
  geom_polygon(data=AllCounty, aes(x=long, y=lat, group=group),
               color="gray", fill="white", size = .1 ) +
  geom_polygon(data = cTable, aes(x = long, y = lat, group = group, 
                                  fill = `projectAmount > 100000000`),
                color = "grey", size = 0.2, alpha = 1.6) +
  geom_polygon(data = MainStates, aes(x = long, y = lat, group = group),
               color="black", fill="white", size = 0.2, alpha = 0.3) +
  scale_fill_manual(values = c("darkblue", "red"))+ 
  ggtitle("County Based TOtal Project Amount") +
  
  # Center the title
  theme(plot.title = element_text(hjust = 0.5))

```






